---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r data}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And lets preview this data:

```{r head}
head(inc)
```
```{r summary}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r nonvis1}
# Insert your code here, create more chunks as necessary
tail(inc)
```
```{r vis2}
# dimensions
dim(inc)
```
```{r vis3}
# column names
colnames(inc)

```
```{r vis4}
# total nas
sum(is.na(inc))
```
## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r question1}
# Answer Question 1 here

plot(table(inc$State))

```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r question 2}
# Answer Question 2 here

# Sort and Filter top 3
state_max = as.matrix(table(inc$State))

third_largest_state = state_max[order(state_max[,1],decreasing=TRUE),][3]
print(third_largest_state)

filt_crit = "NY"

third_largest_filt_data = subset(inc,State == filt_crit)
third_largest_filt_data = third_largest_filt_data[complete.cases(third_largest_filt_data),]
head(third_largest_filt_data)


## initial Plots

mean_third_data = aggregate(third_largest_filt_data$Employees,list(third_largest_filt_data$Industry),FUN=mean)

median_third_data = aggregate(third_largest_filt_data$Employees,list(third_largest_filt_data$Industry),FUN=median)

mean_third_data$Group.1 = as.factor(mean_third_data$Group.1)
median_third_data$Group.1 = as.factor(median_third_data$Group.1)

par(cex.lab=1,cex.axis=.35)
plot(mean_third_data$Group.1,mean_third_data$x,las=2,ylab='Employees',main = 'Mean Employees NY',xlab="")
plot(median_third_data$Group.1,median_third_data$x,las=2,ylab='Employees',main = 'Median Employees NY',xlab="")

## check outlier
boxplot(third_largest_filt_data[third_largest_filt_data$Industry=='Business Products & Services',]$Employees)
boxplot(third_largest_filt_data[third_largest_filt_data$Industry=='Consumer Products & Services',]$Employees)
boxplot(third_largest_filt_data[third_largest_filt_data$Industry=='Human Resources',]$Employees)
boxplot(third_largest_filt_data[third_largest_filt_data$Industry=='Travel & Hospitality',]$Employees)

boxplot(third_largest_filt_data[third_largest_filt_data$Industry=='Energy',]$Employees)

boxplot(third_largest_filt_data[third_largest_filt_data$Industry=='Environmental Services',]$Employees)

## Remove Outlier

new_filt_table = subset(third_largest_filt_data,Rank!=4577 & Rank!=4936 & Rank!=1499 & Rank!=2995 & Rank!=3136 & Rank!=3899 & Rank!=4003 & Rank!=4747 & Rank!=2556 & Rank!=2675)


newmean_third_data = aggregate(new_filt_table$Employees,list(new_filt_table$Industry),FUN=mean)

newmedian_third_data = aggregate(new_filt_table$Employees,list(new_filt_table$Industry),FUN=median)

newmean_third_data$Group.1 = as.factor(newmean_third_data$Group.1)
newmedian_third_data$Group.1 = as.factor(newmedian_third_data$Group.1)

par(cex.lab=1,cex.axis=.35)
plot(newmean_third_data$Group.1,newmean_third_data$x,las=2,ylab='Employees',main = 'Mean Employees NY',xlab="")
plot(newmedian_third_data$Group.1,newmedian_third_data$x,las=2,ylab='Employees',main = 'Median Employees NY',xlab="")



```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
# Answer Question 3 here


agg_data_rev_emp = aggregate(list(emp=third_largest_filt_data$Employees,rev=third_largest_filt_data$Revenue),list(industry=third_largest_filt_data$Industry),FUN=sum)

agg_data_rev_emp$rev_per_emp = agg_data_rev_emp$rev/agg_data_rev_emp$emp
#divide by 1000 for scaling
agg_data_rev_emp$rev_per_emp = agg_data_rev_emp$rev_per_emp/1000

agg_data_rev_emp$industry = as.factor(agg_data_rev_emp$industry)


par(cex.lab=1,cex.axis=.35)
plot(agg_data_rev_emp$industry,agg_data_rev_emp$rev_per_emp,las=2,ylab='Revenue Per Emp (by 1000s)',main = 'Revenue per Emp: Industry',xlab="")

```
